\chapter*{Acknowledgements}

I am deeply grateful to my advisor Prof.\ Matteo Riondato. I remember working as
his teaching assistant in my junior year and asking if I could do research with
him. To my surprise, he said yes without hesitation. Ever since then, he has
been there to support me. When I had rough weeks and showed up with nothing
done, he was patient and understanding. When things were not working as planned,
he managed to instill positivity. When I made mistakes, he continued to place
his trust in me. I am always so inspired by his brilliance and passion for
research that I cannot help but leave our meetings with excitement and a smile.
Because of him, I am now more confident in my abilities. He has taught me so
much, perhaps more than he has realized. Most notably, he has taught me what
outstanding mentorship looks like. Matteo has worked tirelessly to help me. And
I know he has done the same for so many around me. In fact, I am confident he
will continue to do so for the students that follow.

I would also like to give my sincere thanks to the computer science faculty,
particularly to those who taught me: Prof.\ Scott Alfeld, Prof.\ Kristen
Gardner, Prof.\ Scott Kaplan, Prof.\ Lyle McGeoch, Prof.\ Will Rosenbaum, and
Prof.\ Lee Spector. The department's professors have my utmost respect for
continuing to do such a phenomenal job in the face of unprecedented
circumstances. I would have never been able to write a thesis without all that I
have learned from them.

After my thesis talk, Prof.\ Lee Spector posed the question of whether
distortion is always undesirable. This question helped bring forth the
realization that some crucial ideas of this work were not emphasized enough. I
hope these ideas are now clear, and I would like to express my appreciation.

Having worked on a similar research topic, Stefan Walzer-Goldfeld and Steedman
Jenkins helped me get up to speed in the beginning and provided feedback on the
thesis later on. Their contributions were imperative, and I could not have asked
for better partners.

Lee Jiaen, Holden Lee, and Sarah Park collaborated with me on multiple computer
science assignments. There were times when we hit our heads against the
wall,\footnote{Mostly figuratively$\ldots$} but there were also times when we
shouted in celebration after things finally worked. They have been part of what
makes computer science fun.

Lesley Zheng was like another older sister to me. She was one of my teaching
assistants when I started taking computer science courses. Since then, she has
been more than happy to help me, whether with the major or my future career. I
hope that I can return the favor someday.

Friends from the Cohan suite, the Greenway suites, and my international family
are responsible for so much of the joy in my life. Each one of them has made
Amherst more memorable, and for that, I feel extremely blessed. I know they will
all achieve great things, and I will miss them a lot.

I cannot talk about my time at Amherst without mentioning Sona. She has been so
good to me. Whenever I felt overwhelmed, she gave me comfort. Whenever I was
working too much, she forced me to live a little. I only hope that she feels
supported too, as I am so grateful for everything she has done.

Jason Liu, Hiro Wang, Joanne Wang, and Sean Wang grew up with me in Beijing, and
we moved to the U.S.\ for college around the same time. Even though I have not
been able to return to Beijing in a while, seeing them has always brought a
sense of home and comfort. I am so thankful for our friendship.

Ariana overlapped with me at Amherst for two years. I am so lucky to have her as
an older sister since I would be completely lost without her. I somehow manage
to follow her wherever she goes, so while we are living apart now, I have a
feeling we will be with each other soon.

I am greatly indebted to my parents. They have continually given me so much love
and support from across the world, even during times when I took them for
granted. I hope one day I can give back to them even half of what I have
received, as they have given me so much.
