%! TEX root = thesis.tex
% Keep the list in alphabetical order, unless is really helpful to break it =)
\usepackage{mathtools} % for DeclarePairedDelimiter
\usepackage{accents} % to use custom accents
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath} % additional-but-essentially-standard mathematical commands
\usepackage{amssymb} % additional-but-essentially-standard mathematical symbols
\usepackage{amsthm} % additional-but-essentially-standard mathematical theorems
\usepackage{booktabs} % better tables
\usepackage{dsfont} % for font for indicator function
\usepackage{natbib} % for bibliography
\usepackage{graphicx} % package for images
\graphicspath{ {./images/} } % path for images
\usepackage[unicode]{hyperref} % reference links. Must be loaded before cleveref
\hypersetup{pdfborder = {0 0 0}} % do not put a border around links
\usepackage{multirow} % for multi-line rows in tables
\usepackage{tikz} % for drawing a custom accent
\usetikzlibrary{decorations.pathmorphing}
\usepackage{xfrac} % for nice inline fractions
\usepackage[font=small,labelfont=bf,labelsep=space]{caption}
\usepackage[font=small,labelfont=bf]{subcaption} % for subfigures
\usepackage{cleveref}
\crefname{algorithm}{Alg.}{Algs.}
\Crefname{algorithm}{Algorithm}{Algorithms}
\crefname{appendix}{App.}{App.}
\Crefname{appendix}{Appendix}{Appendices}
\crefname{chapter}{Ch.}{Ch.}
\Crefname{chapter}{Chapter}{Chapters}
\crefname{corollary}{Corol.}{Corolls.}
\Crefname{corollary}{Corollary}{Corollaries}
\crefname{conjecture}{Conjecture}{Conjectures}
\Crefname{conjecture}{Conjecture}{Conjectures}
\crefformat{conjecture}{Conjecture~#2#1#3}
\crefmultiformat{conjecture}{Conjectures~#2#1#3}{\ and~#2#1#3}{, #2#1#3}{\ and~#2#1#3}
\crefname{definition}{Def.}{Defs.}
\Crefname{definition}{Definition}{Definition}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Figure}{Figures}
\crefname{lemma}{Lemma}{Lemmas}
\Crefname{lemma}{Lemma}{Lemmas}
\crefname{proposition}{Prop.}{Props.}
\Crefname{proposition}{Proposition}{Propositions}
\Crefname{section}{Section}{Sections}
\crefname{section}{Sect.}{Sect.}
\crefname{subsection}{Sect.}{Sect.}
\Crefname{subsection}{Section}{Sections}
\crefname{subsubsection}{Sect.}{Sect.}
\Crefname{subsubsection}{Section}{Sections}
\crefname{table}{Table}{Tables}
\Crefname{table}{Table}{Tables}
\crefname{theorem}{Thm.}{Thms.}
\Crefname{theorem}{Theorem}{Theorems}

\usepackage{chato-notes} % temporary package. Remove/comment out before submission
