%! TEX root = diffusr.tex

\begin{thebibliography}{37}

%%% ====================================================================
%%% NOTE TO THE USER: you can override these defaults by providing
%%% customized versions of any of these macros before the \bibliography
%%% command.  Each of them MUST provide its own final punctuation,
%%% except for \shownote{}, \showDOI{}, and \showURL{}.  The latter two
%%% do not use final punctuation, in order to avoid confusing it with
%%% the Web address.
%%%
%%% To suppress output of a particular field, define its macro to expand
%%% to an empty string, or better, \unskip, like this:
%%%
%%% \newcommand{\showDOI}[1]{\unskip}   % LaTeX syntax
%%%
%%% \def \showDOI #1{\unskip}           % plain TeX syntax
%%%
%%% ====================================================================

\ifx\showCODEN\undefined\def\showCODEN#1{\unskip}\fi
\ifx\showDOI\undefined\def\showDOI#1{#1}\fi
\ifx\showISBNx\undefined\def\showISBNx#1{\unskip}\fi
\ifx\showISBNxiii\undefined\def\showISBNxiii#1{\unskip}\fi
\ifx\showISSN\undefined\def\showISSN#1{\unskip}\fi
\ifx\showLCCN\undefined\def\showLCCN#1{\unskip}\fi
\ifx\shownote\undefined\def\shownote#1{#1}\fi
\ifx\showarticletitle\undefined\def\showarticletitle#1{#1}\fi
\ifx\showURL\undefined\def\showURL{\relax}\fi
% The following commands are used for tagged output and should be
% invisible to TeX
\providecommand\bibfield[2]{#2}
\providecommand\bibinfo[2]{#2}
\providecommand\natexlab[1]{#1}
\providecommand\showeprint[2][]{arXiv:#2}

\bibitem[Agrawal and Srikant(1994)]%
        {AgrawalS94}
\bibfield{author}{\bibinfo{person}{R. Agrawal} {and}
  \bibinfo{person}{R. Srikant}.} \bibinfo{year}{1994}\natexlab{}.
\newblock\showarticletitle{Fast Algorithms for Mining Association Rules in
  Large Databases}. In VLDB'94.

\bibitem[Bonferroni(1936)]%
        {Bonferroni37}
\bibfield{author}{\bibinfo{person}{C.\ E. Bonferroni}.}
  \bibinfo{year}{1936}\natexlab{}.
\newblock\showarticletitle{Teoria statistica delle classi e calcolo delle
  probabilit{\`a}}.
\newblock\bibinfo{journal}{\emph{Pubb.\ R.\ Inst.\ Sup.\ Sci.\ Econ. Commerc.\
Firenze}}  \bibinfo{volume}{8} (\bibinfo{year}{1936}), \bibinfo{pages}{3--62}.

\bibitem[Brin et~al\mbox{.}(1997)]%
        {BrinMS97}
\bibfield{author}{\bibinfo{person}{S. Brin}, \bibinfo{person}{R.
  Motwani}, {and} \bibinfo{person}{C. Silverstein}.}
  \bibinfo{year}{1997}\natexlab{}.
\newblock\showarticletitle{Beyond market baskets: Generalizing association
  rules to correlations}. In SIGMOD'97.

\bibitem[Duivesteijn and Knobbe(2011)]%
        {DuivesteijnK11}
\bibfield{author}{\bibinfo{person}{W. Duivesteijn} {and}
  \bibinfo{person}{A. Knobbe}.} \bibinfo{year}{2011}\natexlab{}.
\newblock\showarticletitle{Exploiting False Discoveries --- Statistical
  Validation of Patterns and Quality Measures in Subgroup Discovery}. In
  ICDM'11.

\bibitem[Ferkingstad et~al\mbox{.}(2015)]%
        {FerkingstadHS15}
\bibfield{author}{\bibinfo{person}{E. Ferkingstad}, \bibinfo{person}{L. Holden},
  {and} \bibinfo{person}{G. K. Sandve}.}
  \bibinfo{year}{2015}\natexlab{}.
\newblock\showarticletitle{{Monte Carlo} Null Models for Genomic Data}.
\newblock\bibinfo{journal}{\emph{Statist. Sci.}} \bibinfo{volume}{30},
  \bibinfo{number}{1} (\bibinfo{year}{2015}), \bibinfo{pages}{59--71}.

\bibitem[Gionis et~al\mbox{.}(2007)]%
        {GionisMMT07}
\bibfield{author}{\bibinfo{person}{A. Gionis}, \bibinfo{person}{H. Mannila},
  \bibinfo{person}{T. Mielik{\"a}inen}, {and} \bibinfo{person}{P. Tsaparas}.}
  \bibinfo{year}{2007}\natexlab{}.
\newblock\showarticletitle{Assessing data mining results via swap
  randomization}.
\newblock\bibinfo{journal}{\emph{ACM TKDD}} \bibinfo{volume}{1}, \bibinfo{number}{3}
  (\bibinfo{year}{2007}), \bibinfo{pages}{14}.

\bibitem[G{\"u}nnemann et~al\mbox{.}(2012)]%
        {GunnemannDJE12}
\bibfield{author}{\bibinfo{person}{S. G{\"u}nnemann},
  \bibinfo{person}{P. Dao}, \bibinfo{person}{M. Jamali}, {and}
  \bibinfo{person}{M. Ester}.} \bibinfo{year}{2012}\natexlab{}.
\newblock\showarticletitle{Assessing the Significance of Data Mining Results
  on Graphs with Feature Vectors}. In ICDM'12.

\bibitem[H{\"a}m{\"a}l{\"a}inen(2010)]%
        {Hamalainen10}
\bibfield{author}{\bibinfo{person}{W. H{\"a}m{\"a}l{\"a}inen}.}
  \bibinfo{year}{2010}\natexlab{}.
\newblock\showarticletitle{{StatApriori}: an efficient algorithm for searching
  statistically significant association rules}.
\newblock\bibinfo{journal}{\emph{KAIS}}
  \bibinfo{volume}{23}, \bibinfo{number}{3} (\bibinfo{year}{2010}),
  \bibinfo{pages}{373--399}.

\bibitem[H{\"a}m{\"a}l{\"a}inen and Webb(2019)]%
        {HamalainenW19}
\bibfield{author}{\bibinfo{person}{W. H{\"a}m{\"a}l{\"a}inen} {and}
  \bibinfo{person}{G.\ I. Webb}.} \bibinfo{year}{2019}\natexlab{}.
\newblock\showarticletitle{A tutorial on statistically sound pattern
  discovery}.
\newblock\bibinfo{journal}{\emph{DMKD}}
  \bibinfo{volume}{33}, \bibinfo{number}{2} (\bibinfo{year}{2019}),
  \bibinfo{pages}{325--377}.

\bibitem[Hanhij{\"a}rvi(2011)]%
        {Hanhijarvi11}
\bibfield{author}{\bibinfo{person}{S. Hanhij{\"a}rvi}.}
  \bibinfo{year}{2011}\natexlab{}.
\newblock\showarticletitle{Multiple hypothesis testing in pattern discovery}.
  In DS'11.

\bibitem[Hanhij{\"a}rvi et~al\mbox{.}(2009)]%
        {HanhijarviGP09}
\bibfield{author}{\bibinfo{person}{S. Hanhij{\"a}rvi},
  \bibinfo{person}{G.\ C. Garriga}, {and} \bibinfo{person}{K. Puolam{\"a}ki}.}
  \bibinfo{year}{2009}\natexlab{}.
\newblock\showarticletitle{Randomization Techniques for Graphs}. In SDM'09.


\bibitem[Hanhij\"{a}rvi et~al\mbox{.}(2009)]%
        {HanhijarviOVPTM09}
\bibfield{author}{\bibinfo{person}{S. Hanhij\"{a}rvi},
  \bibinfo{person}{M. Ojala}, \bibinfo{person}{N. Vuokko},
  \bibinfo{person}{K. Puolam\"{a}ki}, \bibinfo{person}{N. Tatti}, {and}
  \bibinfo{person}{H. Mannila}.} \bibinfo{year}{2009}\natexlab{}.
\newblock\showarticletitle{Tell Me Something {I} Don't Know: Randomization
  Strategies for Iterative Data Mining}. In KDD'09.

\bibitem[Holm(1979)]%
        {Holm79}
\bibfield{author}{\bibinfo{person}{S. Holm}.}
  \bibinfo{year}{1979}\natexlab{}.
\newblock\showarticletitle{A simple sequentially rejective multiple test
  procedure}.
\newblock\bibinfo{journal}{\emph{Scand.\ J. Stat.}}
  \bibinfo{volume}{6}, \bibinfo{number}{2} (\bibinfo{year}{1979}),
  \bibinfo{pages}{65--70}.

\bibitem[Kirsch et~al\mbox{.}(2012)]%
        {KirschMPPUV12}
\bibfield{author}{\bibinfo{person}{A. Kirsch}, \bibinfo{person}{M.
  Mitzenmacher}, \bibinfo{person}{A. Pietracaprina},
  \bibinfo{person}{G. Pucci}, \bibinfo{person}{E. Upfal}, {and}
  \bibinfo{person}{F. Vandin}.} \bibinfo{year}{2012}\natexlab{}.
\newblock\showarticletitle{An efficient rigorous approach for identifying
  statistically significant frequent itemsets}.
\newblock\bibinfo{journal}{\emph{JACM}}
  \bibinfo{volume}{59}, \bibinfo{number}{3} (\bibinfo{year}{2012}),
  \bibinfo{pages}{1--22}.

\bibitem[Komiyama et~al\mbox{.}(2017)]%
        {KomiyamaIANM17}
\bibfield{author}{\bibinfo{person}{J. Komiyama}, \bibinfo{person}{M.
  Ishihata}, \bibinfo{person}{H. Arimura}, \bibinfo{person}{T.
  Nishibayashi}, {and} \bibinfo{person}{S.-I. Minato}.}
  \bibinfo{year}{2017}\natexlab{}.
\newblock\showarticletitle{Statistical Emerging Pattern Mining with Multiple
  Testing Correction}. In KDD'17.

\bibitem[Kontonasios et~al\mbox{.}(2012)]%
        {KontonasiosSDB12}
\bibfield{author}{\bibinfo{person}{K.-N. Kontonasios},
  \bibinfo{person}{E. Spyropoulou}, {and} \bibinfo{person}{T. De~Bie}.}
  \bibinfo{year}{2012}\natexlab{}.
\newblock\showarticletitle{Knowledge discovery interestingness measures based
  on unexpectedness}.
\newblock\bibinfo{journal}{\emph{{WIREs} DMKD}}
  \bibinfo{volume}{2}, \bibinfo{number}{5} (\bibinfo{year}{2012}),
  \bibinfo{pages}{386--399}.

\bibitem[Megiddo and Srikant(1998)]%
        {MegiddoS98}
\bibfield{author}{\bibinfo{person}{N. Megiddo} {and}
  \bibinfo{person}{R. Srikant}.} \bibinfo{year}{1998}\natexlab{}.
\newblock\showarticletitle{Discovering predictive association rules}. In
KDD'94.

\bibitem[Llinares-L{\'o}pez et~al\mbox{.}(2015)]%
        {LlinaresLopezSPB15}
\bibfield{author}{\bibinfo{person}{F. Llinares-L{\'o}pez},
  \bibinfo{person}{M.Sugiyama}, \bibinfo{person}{L. Papaxanthos},
  {and} \bibinfo{person}{K. Borgwardt}.} \bibinfo{year}{2015}\natexlab{}.
\newblock\showarticletitle{Fast and memory-efficient significant pattern
  mining via permutation testing}. In KDD'15.

\bibitem[Low-Kam et~al\mbox{.}(2013)]%
        {LowKRKP13}
\bibfield{author}{\bibinfo{person}{C. Low-Kam}, \bibinfo{person}{C.
  Ra{\"\i}ssi}, \bibinfo{person}{M. Kaytoue}, {and} \bibinfo{person}{J.
  Pei}.} \bibinfo{year}{2013}\natexlab{}.
\newblock\showarticletitle{Mining statistically significant sequential
  patterns}. In ICDM'13.

\bibitem[Minato et~al\mbox{.}(2014)]%
        {MinatoUTTS14}
\bibfield{author}{\bibinfo{person}{S.-I. Minato}, \bibinfo{person}{T.
  Uno}, \bibinfo{person}{K. Tsuda}, \bibinfo{person}{A. Terada}, {and}
  \bibinfo{person}{J. Sese}.} \bibinfo{year}{2014}\natexlab{}.
\newblock\showarticletitle{A fast method of statistical assessment for
  combinatorial hypotheses based on frequent itemset enumeration}. In
  ECML'PKDD'14.

\bibitem[Mitzenmacher and Upfal(2005)]%
        {MitzenmacherU05}
\bibfield{author}{\bibinfo{person}{M. Mitzenmacher} {and}
  \bibinfo{person}{E. Upfal}.} \bibinfo{year}{2005}\natexlab{}.
\newblock\bibinfo{booktitle}{\emph{Probability and Computing: Randomized
  Algorithms and Probabilistic Analysis}}.
\newblock\bibinfo{publisher}{Cambridge University Press}.

\bibitem[Ojala(2010)]%
        {Ojala10}
\bibfield{author}{\bibinfo{person}{M. Ojala}.}
  \bibinfo{year}{2010}\natexlab{}.
\newblock\showarticletitle{Assessing Data Mining Results on Matrices with
  Randomization}. In ICDM'10.

\bibitem[Ojala et~al\mbox{.}(2010)]%
        {OjalaGGM10}
\bibfield{author}{\bibinfo{person}{M. Ojala}, \bibinfo{person}{G.\ C.
  Garriga}, \bibinfo{person}{A. Gionis}, {and} \bibinfo{person}{H.
  Mannila}.} \bibinfo{year}{2010}\natexlab{}.
\newblock\showarticletitle{Evaluating Query Result Significance in Databases
  via Randomizations}. In SDM'10.

\bibitem[Ojala et~al\mbox{.}(2008)]%
        {OjalaVKHM08}
\bibfield{author}{\bibinfo{person}{M. Ojala}, \bibinfo{person}{N.
  Vuokko}, \bibinfo{person}{A. Kallio}, \bibinfo{person}{N. Haiminen},
  {and} \bibinfo{person}{H. Mannila}.} \bibinfo{year}{2008}\natexlab{}.
\newblock\showarticletitle{Randomization of real-valued matrices for assessing
  the significance of data mining results}. In SDM'08.

\bibitem[Pellegrina et~al\mbox{.}(2019a)]%
        {PellegrinaRV19b}
\bibfield{author}{\bibinfo{person}{L. Pellegrina},
  \bibinfo{person}{M. Riondato}, {and} \bibinfo{person}{F. Vandin}.}
  \bibinfo{year}{2019}\natexlab{a}.
\newblock\showarticletitle{Hypothesis Testing and Statistically-sound Pattern
  Mining}. In KDD'19.

\bibitem[Pellegrina et~al\mbox{.}(2019b)]%
        {PellegrinaRV19a}
\bibfield{author}{\bibinfo{person}{L. Pellegrina},
  \bibinfo{person}{M. Riondato}, {and} \bibinfo{person}{F. Vandin}.}
  \bibinfo{year}{2019}\natexlab{b}.
\newblock\showarticletitle{{SPuManTE}: Significant Pattern Mining with
  Unconditional Testing}. In KDD'19.

\bibitem[Pellegrina and Vandin(2020)]%
        {PellegrinaV20}
\bibfield{author}{\bibinfo{person}{L. Pellegrina} {and}
  \bibinfo{person}{F. Vandin}.} \bibinfo{year}{2020}\natexlab{}.
\newblock\showarticletitle{Efficient mining of the most significant patterns
  with permutation testing}.
\newblock\bibinfo{journal}{\emph{DMKD}}
  \bibinfo{volume}{34} (\bibinfo{year}{2020}), \bibinfo{pages}{1201--1234}.

\bibitem[Stanley(2011)]%
        {Stanley11}
\bibfield{author}{\bibinfo{person}{R.~P. Stanley}.}
  \bibinfo{year}{2011}\natexlab{}.
\newblock\bibinfo{booktitle}{\emph{Enumerative Combinatorics}
  (\bibinfo{edition}{2} ed.)}. Vol.~\bibinfo{volume}{1}.
\newblock\bibinfo{publisher}{Cambridge University Press}.

\bibitem[Tan et~al\mbox{.}(2004)]%
        {TanKS04}
\bibfield{author}{\bibinfo{person}{P.-N. Tan}, \bibinfo{person}{V.
  Kumar}, {and} \bibinfo{person}{J. Srivastava}.}
  \bibinfo{year}{2004}\natexlab{}.
\newblock\showarticletitle{Selecting the right objective measure for
  association analysis}.
\newblock\bibinfo{journal}{\emph{Inf. Sys.}}  \bibinfo{volume}{29}
  (\bibinfo{year}{2004}), \bibinfo{pages}{293--313}.

\bibitem[Terada et~al\mbox{.}(2013)]%
        {TeradaOHTS13}
\bibfield{author}{\bibinfo{person}{A. Terada}, \bibinfo{person}{M.
  Okada-Hatakeyama}, \bibinfo{person}{K. Tsuda}, {and} \bibinfo{person}{J.
  Sese}.} \bibinfo{year}{2013}\natexlab{}.
\newblock\showarticletitle{Statistical significance of combinatorial
  regulations}.
\newblock\bibinfo{journal}{\emph{PNAS}} \bibinfo{volume}{110}, \bibinfo{number}{32}
  (\bibinfo{year}{2013}), \bibinfo{pages}{12996--13001}.

\bibitem[Tonon and Vandin(2019)]%
        {TononV19}
\bibfield{author}{\bibinfo{person}{A. Tonon} {and} \bibinfo{person}{F.
  Vandin}.} \bibinfo{year}{2019}\natexlab{}.
\newblock\showarticletitle{Permutation strategies for mining significant
  sequential patterns}. In ICDM'19.

\bibitem[Vreeken and Tatti(2014)]%
        {VreekenT14}
\bibfield{author}{\bibinfo{person}{J. Vreeken} {and}
  \bibinfo{person}{N. Tatti}.} \bibinfo{year}{2014}\natexlab{}.
\newblock\showarticletitle{Interesting patterns}.
\newblock%
In \bibinfo{booktitle}{\emph{Frequent pattern mining}}.
  \bibinfo{publisher}{Springer}, \bibinfo{pages}{105--134}.

\bibitem[Wasserman(2005)]%
        {Wasserman05}
\bibfield{author}{\bibinfo{person}{L. Wasserman}.}
  \bibinfo{year}{2005}\natexlab{}.
\newblock\bibinfo{booktitle}{\emph{All of Statistics: A Concise Course in
  Statistical Inference}}.
\newblock\bibinfo{publisher}{Springer}.

\bibitem[Webb(2006)]%
        {Webb06}
\bibfield{author}{\bibinfo{person}{G.\ I. Webb}.}
  \bibinfo{year}{2006}\natexlab{}.
\newblock\showarticletitle{Discovering significant rules}. In KDD'06.

\bibitem[Webb(2007)]%
        {Webb07}
\bibfield{author}{\bibinfo{person}{G.\ I. Webb}.}
  \bibinfo{year}{2007}\natexlab{}.
\newblock\showarticletitle{Discovering significant patterns}.
\newblock\bibinfo{journal}{\emph{Mach.\ Learn.}} \bibinfo{volume}{68},
  \bibinfo{number}{1} (\bibinfo{year}{2007}), \bibinfo{pages}{1--33}.

\bibitem[Webb(2008)]%
        {Webb08}
\bibfield{author}{\bibinfo{person}{G.\ I Webb}.}
  \bibinfo{year}{2008}\natexlab{}.
\newblock\showarticletitle{Layered critical values: a powerful
  direct-adjustment approach to discovering significant patterns}.
\newblock\bibinfo{journal}{\emph{Mach.\ Learn.}} \bibinfo{volume}{71},
  \bibinfo{number}{2--3} (\bibinfo{year}{2008}), \bibinfo{pages}{307--323}.

\bibitem[Westfall and Young(1993)]%
        {WestfallY93}
\bibfield{author}{\bibinfo{person}{P.\ H. Westfall} {and}
  \bibinfo{person}{S.\ S. Young}.} \bibinfo{year}{1993}\natexlab{}.
\newblock\bibinfo{booktitle}{\emph{Resampling-Based Multiple Testing: Examples
  and Methods for {p}-Value Adjustment}}.
\newblock\bibinfo{publisher}{Wiley-Interscience}.

\bibitem[Wu et~al\mbox{.}(2016)]%
        {WuHGLZY16}
\bibfield{author}{\bibinfo{person}{J. Wu}, \bibinfo{person}{Z. He},
  \bibinfo{person}{F. Gu}, \bibinfo{person}{X. Liu},
  \bibinfo{person}{J. Zhou}, {and} \bibinfo{person}{C. Yang}.}
  \bibinfo{year}{2016}\natexlab{}.
\newblock\showarticletitle{Computing exact permutation {$p$}-values for
  association rules}.
\newblock\bibinfo{journal}{\emph{Inf.\ Sci.}}  \bibinfo{volume}{346}
  (\bibinfo{year}{2016}), \bibinfo{pages}{146--162}.

\end{thebibliography}
